---
title: "PS270 Final Project"
author: "Ella Mahaney"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts



1) My research question is how does the month impact the pressure at the center of a hurricane in the Atlantic?
My hypothesis is in month 9 when there are the most hurricanes the pressure in the center of those hurricanes will be the strongest. I will be looking at how many hurricanes occured in month 9 at 1014 millibars of presure. I chose 1014 as my measure of high pressure because it is near the highest pressure recorded in the data set. This will be the case because since there are the most hurricanes during this month it is optimal conditions and leads to stronger hurricanes. I think this is interesting because the pressure may help indicate when a hurricane will occur. This can help people know when it is safe to travel to areas where there are typically hurricanes.

2) 
  The data source I chose features information about hurricanes from 1975 to 2021. This data lists the months that hurricanes occur along with how many there were in that month. This data is recorded by NOAA the national Ocean and Atmospheric Administration.
  The independent variable is measured based on the date. However the independent variable is measured based on coded amount of hurricanes per month and pressure levels. This plot showcases the dependent variable pressure at the center of hurricanes.

```{r}
file.exists("data/storms.csv")
```

```{r}
storms_data <- read.csv("data/storms.csv")

```

```{r}
library(tidyverse)
library(knitr)

new_storms_data <- storms_data |> 
  filter(month == 9)

new_storms_data |>
  ggplot(data = storms,
       mapping = aes(x = pressure)) + 
  geom_bar() + 
  labs(
    title = "Number of Hurricanes that Occur each month",
    x = "Pressure of Hurricane",
    y = "Amount of Hurricanes Recorded"
  ) 


```

3)
```{r}
library(tidyverse)
library(knitr)
storms_data |>
  ggplot(data = storms,
       mapping = aes(x = month)) + 
  geom_bar() + 
  labs(
    title = "Number of Hurricanes that Occur each month",
    x = "Month that Hurricane Occurs",
    y = "Amount of Hurricanes Recorded"
  ) 


```
```{r}
library(infer)
storms_data <- storms_data |> 
  mutate(new_month = if_else(month == 9, "True", "False") |> as.factor(), new_pressure = if_else(pressure == 1014, "High", "Low") |> as.factor())
  


regression <- storms_data |> 
  specify(new_pressure ~ new_month, success = "High") |> 
  calculate(stat = "diff in props", order = c("True", "False"))

regression |> 
knitr::kable(digits = 2)
```
  
```{r}
diff_storms_data <- storms_data |> 
specify(new_pressure ~ new_month, success = "High") |> 
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |> 
  calculate(stat = "diff in props", order = c("True", "False"))


  
```
```{r}
diff_storms_data |> 
  visualize() + 
  shade_p_value(obs_stat = regression, direction = "two sided") + 
  labs(
    title = "Difference in Proportion of High vs Low pressure in September",
    x = "Proportion of High Pressure Storms",
    y = "Amount"
  ) 
```


```{r}
base_p <- diff_storms_data |>
  get_p_value(obs_stat = regression, direction = "both")

base_p |> 
knitr::kable(digits = 2)
```

  
  What the main coefficient of difference in terms of difference in preportions means is that since it is a negative number non September months have a higher proportion of high pressure storms occuring. The overall proportion of high pressure storms is 0.17 percent lower in the month of September. 
 
  
  
  What the P value means is that these findings are not statistically significant. This is because the p value 0.202 is bigger than 0.05 which is too large of a p value to be statistically significant. There is a 20.2 percent chance that the difference is due to random variation in the sample. The null hypothesis is not rejected. We should not interpret this relationship as causal. There is no relationship between these variables. 
 

4) 
The overall results are that storms that occur during the month of September (when the most storms are present) do not have the highest pressure levels at their center. Tharefore my hypothesis is not supported. However there are limitations to this analysis the findings were not statistically significant. This means that there is a large margin for error 20% where alterior variables may have effected results. Tharefor it is hard to say the results I found are true. If I had more time I would purchase better equipment to measure pressure levels to ensure they are accurate. I would also hire researchers to comb through the data recorded to ensure no hurricanes were left out. 




https://www.kaggle.com/datasets/utkarshx27/noaa-atlantic-hurricane-database?resource=download





















