---
title: "PS270 Final Project"
author: "Ella Mahaney"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to weather and its impact on people's mood. I think this could be interesting and the results may be relatable to many people.

How does the month impact the pressure at the time of a hurricane?
The later the month is (between the months of 6-10) the higher the pressure at the time of a hurricane. The reason why this happens is because peak hurricane time is around late september early october.
The explanatory variable of interest is the month that it is when a specific hurricane occurs. It is measured based on the date of when the hurricane starts.
The outcome variable is the pressure of the center of the hurricane. It is measured in milibars.
What would support it is if as the pressure of the hurricane increased at the same rate of the months changing. What would disprove it is as the months passed the pressure of a hurricane decreased too.
https://www.kaggle.com/datasets/utkarshx27/noaa-atlantic-hurricane-database?resource=download


Discussion
Overall there was a correlation found between pressure and the month that the hurricane occured. As it became later in the year starting in June the pressure decreased until October when there were less hurricanes and the pressure increased.
```{r}
file.exists("data/storms.csv")
```

```{r}
storms_data <- read.csv("data/storms.csv")
storms_data
```
```{r}
library(tidyverse)
library(knitr)
storms_data |>
  ggplot(data = storms,
       mapping = aes(x = month)) + 
  geom_bar() + 
  labs(
    title = "Number of Hurricanes that Occur each month",
    x = "Month that Hurricane Occurs",
    y = "Amount of Hurricanes Recorded"
  ) 


```


```{r}
outcome_month_7 <- storms_data |> 
  filter(month == 7) |> 
  summarize(
    month_7_mean = mean(pressure, na.rm = TRUE)
  )

outcome_month_7 |>
knitr::kable(caption = "Month 7 Pressure Average")
```





```{r}


outcome_month_8 <- storms_data |> 
  filter(month == 8) |> 
  summarize(
    month_8_mean = mean(pressure, na.rm = TRUE)
  )
outcome_month_8 |>
knitr::kable(caption = "Month 8 Pressure Average")
```
```{r}
outcome_month_9 <- storms_data |> 
  filter(month == 9) |> 
  summarize(
    month_9_mean = mean(pressure, na.rm = TRUE)
  )
outcome_month_9 |>
knitr::kable(caption = "Month 9 Pressure Average")
```

```{r}
outcome_month_10 <- storms_data |> 
  filter(month == 10) |> 
  summarize(
    month_10_mean = mean(pressure, na.rm = TRUE)
  )
outcome_month_10 |>
knitr::kable(caption = "Month 10 Pressure Average")
```

